---
title: "Final Project - McDonalds"
author: "Andrew Tran"
date: "May 11, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

  McDonald's is the world's largest restaurant chain by revenue, serving over 69 million customers daily in over 100 countries. It is the stereotypical ideal that comes to mind when one thinks of fast food, known for its hamburgers, fries, soft drinks, and other menu items. In today's current culture, with an increased focus on having a healthy lifestyle, people often disdain McDonald's for its supposed unhealtiness, its food and drink loaded with fat, grease, and sugar. But does McDonald's really deserve such a bad reputation? 
  
  As a person who has eaten at the UMD Stamp McDonalds an untold number of times, I decided to create this tutorial on exploring the nutrition behind the menu items sold at McDonald's. 
  

## Part 1 - Data Loading and Cleaning 
  
  The very first thing we must do is to retrieve the data we wish to look at and analyze. In this case, we will be using a free dataset provided by Kaggle: 
  
https://www.kaggle.com/mcdonalds/nutrition-facts

  Once we've downloaded the dataset to our computer, let's load it up into R-Studio, along with a few basic libraries we will need. 
  
  Of course, if you want to look at the official information as provided by McDonald's, you are always free to visit their official nutritional information calculator for yourself:
  
https://www.mcdonalds.com/us/en-us/about-our-food/nutrition-calculator.html

```{r loading_data}
library(tidyverse)
library(ggplot2)
library(dplyr)


# Use the read_csv function to load the dataset into a variable we can use. 
menu <- read_csv("C:/Users/Andrew/Desktop/College/CMSC320/Final/menu.csv")
menu
```


  Now that our dataset has been loaded in R-Studio as a dataframe, we can see that it contains 260 different McDonald's menu items (entities, or rows), each with 24 different attributes (columns). At first glance, this dataset appears to be very well organized too, so we may not need to do much cleaning with it. Let's use colnames() to see every single attribute used in this dataframe, and find out what we can work with. 

```{r col_names}
colnames(menu)
```


  Looks pretty good so far. There is however, one problem. Several of these attributes are missing their respective units of measurement. "% Daily Values" are easy enough to understand, but what about things like Sodium or Total Fat? What exactly does it mean if an Egg McMuffin possesses a value of 3 for Sugars and a value of 17 for Protein? We can make things clearer than this. 
  
  Let's add the official unit of measurement to their respective units. We can do so by using the rename function along with a pipeline (the "%>%" symbol). The rename function is exactly as it sounds, allowing us to change the names of certain attributes. The pipeline 
  

  
```{r add_units}
# Note that since our attribute names possess spaces, we will need to use backtabs `` in order to access them.
# For example: `attribute name` 
menu <- menu %>%
  rename( `Total Fat (g)`=`Total Fat`, 
         `Saturated Fat (g)`=`Saturated Fat`, 
         `Trans Fat (g)`=`Trans Fat`,
         `Cholesterol (mg)`=`Cholesterol`,
         `Sodium (mg)`=`Sodium`,
         `Carbohydrates (g)`=`Carbohydrates`,
         `Dietary Fiber (g)`=`Dietary Fiber`,
         `Sugars (g)`=`Sugars`,
         `Protein (g)`=`Protein`)

colnames(menu)
```
  
  
## Part 2 - Exploratory Data Analysis (EDA)

  Now that we have our data, let try to do some analyzing, and see if we can find a few things out or two.
  
 Let's start with something simple. For most people, the most common criticisms of fast food are high fat content, sodium, calorie count, and sugar. Let's try finding out the average values for these attributes for all food categories offered by McDonald's. 
  
  To do so, we can use the summarize and group_by functions. Group_by allows use to group these menu items by a certain attribute, while summarize allows us to collapse the dataframe into a single row containing the desired attribute summaries we want (e.g. mean, median, max, min, etc.)
  
```{r menu_select}
menu_averages <- menu %>%
  group_by(Category) %>%
  summarize(`Average Calories`=mean(Calories),
            `Average Sodium (mg)`=mean(`Sodium (mg)`),
            `Average Total Fat (g)` = mean(`Total Fat (g)`),
            `Average Sugars (g)` = mean(`Sugars (g)`)) 
           

menu_averages 
```


  Oho. These numbers do appear to be quite high. But what exactly do they mean? Perhaps we are putting too much focus on raw values at this point. To understand and go one step further, let's put this in context with the % Daily Values required of a normal 2000 Calorie diet. After all, a seemingly high amount of fat, sodium, sugar may actually mean nothing if it is quit low in terms of your actual daily needs. 
  
  Let us again use summarize and group_by to find out our wanted percentages. One thing is missing from our original dataset however: the % daily value in regards to calories per menu item. We can add this in using the mutate function, which adds new attributes to datasets based on the result of an expression. Let us again assume we are using a 2000 Calorie Diet, as specified on today's current nutrition labels. 
  
```{r add_daily_calorie_percentage}
menu <- menu %>%
  # mutate is used to add new attributes based on the result of a given expression 
  mutate(`Calories (% Daily Value)` = as.integer((Calories/2000) * 100))

# Use summarize and group_by to obtain the average daily values we want for each item category 
menu_averages <- menu %>%
  group_by(Category) %>%
  summarize(`Average Calories (% Daily Value)`=mean(`Calories (% Daily Value)`),
            `Average Sodium (% Daily Value)`=mean(`Sodium (% Daily Value)`),
            `Average Total Fat (% Daily Value)` = mean(`Total Fat (% Daily Value)`),
            # In terms of nutrition, sugar falls under carbohydrates for % daily values 
            `Average Carbohydrates (% Daily Value)` = mean(`Carbohydrates (% Daily Value)`)) 
```
 
  Let's try putting our newfound data onto plots, so we can more easily visualize our findings. To do so, we will be using the ggplot package. Plotting, or more formally "data visualization" is an extremely useful tool for any data scientist, allowing us to easily see trends or insights into dataset characteristics. 
  
  ggplot is an extremely versatile and useful package, designed to work well with the tidyverse set of packages. For more information and resources on this package, look here: 
  
  http://ggplot2.tidyverse.org/
  https://www.rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf
  
  
```{r plot_average_daily_values}
menu_averages %>%
  ggplot(mapping=aes(x=Category, y=`Average Calories (% Daily Value)`)) +
    geom_bar(aes(fill=Category), stat="identity") + ggtitle("Average Amount of Calories (% Daily Value) vs. Menu Category") + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) 

# plot mean sodium 
menu_averages %>%
  ggplot(mapping=aes(x=Category, y=`Average Sodium (% Daily Value)`)) +
    geom_bar(aes(fill=Category), stat="identity") + ggtitle("Average Amount of Sodium (% Daily Value) vs. Menu Category") + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) 

# plot mean total fat  
menu_averages %>%
  ggplot(mapping=aes(x=Category, y=`Average Total Fat (% Daily Value)`)) +
    geom_bar(aes(fill=Category), stat="identity") + ggtitle("Average Amount of Total Fat (% Daily Value) vs. Menu Category") + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) 

# plot mean sugars
menu_averages %>%
  ggplot(mapping=aes(x=Category, y=`Average Carbohydrates (% Daily Value)`)) +
    geom_bar(aes(fill=Category), stat="identity") + ggtitle("Average Amount of Carbohydrates (% Daily Value) vs. Menu Category") + theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) 
```

Well now, this certainly isn't looking too good. We can see that Beef & Pork, Breakfast, and Chicken & Fish category menu items easily break over 40 % of one's daily sodium intake. Likewise, they hover around 40% of a person's daily intake of fat. Somewhat unsurprisingly, Smoothies & Shakes are also extremely high in sugar, averaging around 30% of one's daily recommended amount of sugar. Although, it is interesting to note that they easily outstrip Desserts in terms of % Daily Value for Sugar in a ratio of almost 3:1. Perhaps McDonald's really does deserve its reputation as unhealthy, salty, sugary, fatty goodness. 



 Well, with that concluded, let us now ask another interesting question: How much of each nutrient directly contributes to Calories for McDonald's items? That is, what eactly comprises the Calorie count of your average McDonald's menu item? Let's find out. 









   